---
title: "Project"
author: "Marjorie Blanco"
date: "February 7, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(readxl)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse)
library(car)
library(psych)
```

```{r, warning=FALSE, include=FALSE}
getskewness <- function(v) {
   3 * (mean(v) - median(v)) / sd(v)
}

plotHistFunc <- function(x, f = "dodgerblue", na.rm = TRUE, ...) {
  nm <- names(x)
  for (i in seq_along(nm)) {
    p <- ggplot(x,aes_string(x = nm[i])) + 
      geom_histogram(alpha = .5, fill = f) +
      scale_x_log10()
    print(p)
  }
}


plotHistIQRFunc <- function(x, f = "dodgerblue", na.rm = TRUE, ...) {
  nm <- names(x)
  for (i in seq_along(nm)) {
    lower <- quantile(x[,i], .25) - (1.5 * IQR(x[,i])) 
    upper <- quantile(x[,i], .75) + (1.5 * IQR(x[,i]))

    p <- ggplot(x,aes_string(x = nm[i])) + 
      geom_histogram(alpha = .5, fill = f) +
      geom_vline(xintercept = upper, linetype="solid", color = "red", size=1.5) +
      geom_vline(xintercept = lower, linetype="solid", color = "red", size=1.5) 
    print(p)
  }
}


plotHistZFunc <- function(x, f = "dodgerblue", na.rm = TRUE, ...) {
  nm <- names(x)
  for (i in seq_along(nm)) {
    print(i)
    lower <- quantile(x[,i], .25) - (1.5 * IQR(x[,i])) 
    upper <- quantile(x[,i], .75) + (1.5 * IQR(x[,i]))

    p <- ggplot(x,aes_string(x = nm[i])) + 
      geom_histogram(alpha = .5, fill = f) +
      geom_vline(xintercept = upper, linetype="solid", color = "red", size=1.5) +
      geom_vline(xintercept = lower, linetype="solid", color = "red", size=1.5) 
    print(p)
  }
}

plotBarFunc <- function(x, na.rm = TRUE, ...) {
  nm <- names(x)
  for (i in seq_along(nm)) {
    p <- ggplot(x,aes_string(x = nm[i])) + 
      geom_bar(alpha = .5,fill = "dodgerblue") +
      theme(axis.text.x = element_text(angle = 90, hjust = 1))
    print(p)
  }
}

getCor <- function(x, val) {
  z <- cor(x)
  zdf <- as.data.frame(as.table(z))
  zdf <- arrange(zdf,desc(Freq))
  print(zdf[zdf$Freq  > val & zdf$Var1 !=zdf$Var2 ,])
}

getX <- function(df, column)
{
  print(length(which(df[, column]!=0))/length(df[,column]))
}

qqplot.data <- function (vec) # argument: vector of numbers
{
  # following four lines from base R's qqline()
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

  d <- data.frame(resids = vec)

  ggplot(d, aes(sample = resids)) + stat_qq() + geom_abline(slope = slope, intercept = int)

}

dataQuality <- function(x)
{
  round(length(which(x!=0))/length(x),5)* 100
}


replaceAverge <- function(df, column) {
  x <- unique(df[df$column == 0, c("SMBL")])
  df2  <- df %>% select(ID, SMBL, column) %>% filter(SMBL %in% x)
  
  df3 <- df2 %>% 
    filter(i != 0) %>% 
    group_by(SMBL) %>% 
    summarise(n = n()) 
  
  for (i in df3[df3$n >= 2, c("SMBL")])
  {
    l1 <- df2 %>% filter(SMBL %in% i & column != 0)
    l2 <- df2 %>% filter(SMBL %in% i & column == 0)

    df[ df$ID %in% l2$ID ,column] <- mean(l2)
  }
  df
}

NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
```

# Part one

## Read data set

```{r read1, warning=FALSE}
AAERDATA <- read_excel("~/R/DataMining/Project/AAERDATAHWPart1.xlsx", skip=1, na="0")

AAERDATA$ID <- seq.int(nrow(AAERDATA))
AAERDATA1 <- AAERDATA[,c(1:6,97:99)]  # categorical
AAERDATA <- AAERDATA[,c(7:96,100)]    # numerical
```

## Identify missing value

Identify missing values and take appropriate action – eliminate, average value, etc. 
Justify your decision.

The missing values from following features were replaced using linear regression:

- Sales

- Closing Price

- Number of employees

Records with missing values were set to 0.  Records with "I" and "C" in numeric variables were set to 0.  Columns that contains with more than 85% zero values were removed.

The rest of the missing values will be replaced using mean.


### Converting data columns to numeric

```{r convert, warning=FALSE}
AAERDATA <- data.frame(lapply(AAERDATA, function(x) { as.numeric(x) }))
```

Continous variables are those with type `numberic`

```{r numeric1}
colnames(AAERDATA)[colnames(AAERDATA) == 'DATA.23'] <- 'DATA23'
colnames(AAERDATA)[colnames(AAERDATA) == 'DATA.24'] <- 'DATA24'
split(names(AAERDATA),sapply(AAERDATA, function(x) paste(class(x), collapse=" ")))
```

### Percetage of zeros

Determine the percentage of zeros for each data column

```{r, warning=FALSE}
AAERDATA[is.na(AAERDATA)] <- 0
data_col <- lapply(AAERDATA, function(x){ length(which(x!=0))/length(x)})
data_col <- data.frame(data_col)
AAERDATA3 <- AAERDATA[,data_col[1,] > 0.85]
```

The following columns were removed for having more than 85% of zeros.

```{r}
colnames(AAERDATA[,data_col[1,] <= 0.85])
```

```{r}
colnames(AAERDATA[,data_col[1,] > 0.85])
```

Determine if any data column contains NAs

```{r nas}
colnames(AAERDATA)[colSums(is.na(AAERDATA)) > 0]
```


```{r}
data_col <- data_col %>%
   rownames_to_column %>% 
   gather(var, value, -rowname) %>% 
   spread(rowname, value)

colnames(data_col) <- c("Variable","Freq")
data_col <- data_col  %>% arrange(desc(Freq)) %>% filter(Freq > .5)

getCor(AAERDATA3[,1:ncol(AAERDATA3)-1], val = 0.8)
```

```{r}
getX(AAERDATA3, "DATA123")
getX(AAERDATA3, "DATA18")
```

## Look for redundancies and errors.

Remove numbers and period in the SMBL column.

```{r smbl_clean}
AAERDATA1$SMBL <- gsub("[0-9]*", "", AAERDATA1$SMBL)
AAERDATA1$SMBL <- gsub("\\.", "", AAERDATA1$SMBL)
```

```{r plot1, warning=FALSE, error=FALSE} 
plotHistFunc(AAERDATA3[,-43])
```

## Variable summary

Identify mean, count, sum, median, standard deviation for – SALES, Price_close, Employees

```{r setup1}
AAERDATA3$SMBL <- AAERDATA1$SMBL
AAERDATA3$yeara <- AAERDATA1$yeara
AAERDATA3$ID <- AAERDATA$ID
```

### Sales

The original data quality for Sales is `r round(length(which(AAERDATA3$DATA12!=0))/length(AAERDATA3$DATA12),3)*100`%.

```{r sales1}
dataQuality(AAERDATA3$DATA12)

x <- unique(AAERDATA3[AAERDATA3$DATA12 == 0, c("SMBL")])
AAERDATA4  <- AAERDATA3 %>% select(ID, SMBL, yeara, DATA12) %>% filter(SMBL %in% x)

AAERDATA5 <- AAERDATA4 %>% 
  filter(DATA12 != 0) %>% 
  group_by(SMBL) %>% 
  summarise(n = n()) 

for (i in AAERDATA5[AAERDATA5$n > 2, c("SMBL")])
{
  l1 <- AAERDATA4 %>% filter(SMBL %in% i & DATA12 != 0)
  l2 <- AAERDATA4 %>% filter(SMBL %in% i & DATA12 == 0)
  fit <- lm(data = l1, DATA12~yeara)
  AAERDATA3[ AAERDATA3$ID %in% l2$ID ,c("DATA12")] <- predict(fit, l2[,c("DATA12","yeara")]) 
}

dataQuality(AAERDATA3$DATA12)
```

The improved data quality for Sales is `r round(length(which(AAERDATA3$DATA12!=0))/length(AAERDATA3$DATA12),3)*100`% after missing values were predicted using linear regression.

Mean: `r mean(AAERDATA3$DATA12)`

Count: `r length(AAERDATA3$DATA12 != 0)`

Sum: `r sum(AAERDATA3$DATA12)`

Median: `r median(AAERDATA3$DATA12)`

Standard: `r sd(AAERDATA3$DATA12)`

```{r sales2, warning=FALSE}
#Original data

#Updated data
sales_mean <- mean(AAERDATA3$DATA12)
sales_count <- length(AAERDATA3$DATA12 != 0)
sales_sum <- sum(AAERDATA3$DATA12)
sales_median <- median(AAERDATA3$DATA12)
sales_sd <- sd(AAERDATA3$DATA12)

sales_mean
sales_count
sales_sum
sales_median
sales_sd

plotHistFunc(data.frame(AAERDATA3$DATA12))
```

### Price at closing

The original data quality for Price at closing is `r round(length(which(AAERDATA3$DATA24!=0))/length(AAERDATA3$DATA24),3)*100`%.

```{r price1}
dataQuality(AAERDATA3$DATA24)

x <- unique(AAERDATA3[AAERDATA3$DATA24 == 0, c("SMBL")])
AAERDATA4  <- AAERDATA3 %>% select(ID, SMBL, yeara, DATA24) %>% filter(SMBL %in% x)

AAERDATA5 <- AAERDATA4 %>% 
  filter(DATA24 != 0) %>% 
  group_by(SMBL) %>% 
  summarise(n = n()) 

for (i in AAERDATA5[AAERDATA5$n > 2, c("SMBL")])
{
  l1 <- AAERDATA4 %>% filter(SMBL %in% i & DATA24 != 0)
  l2 <- AAERDATA4 %>% filter(SMBL %in% i & DATA24 == 0)
  fit <- lm(data = l1, DATA24~yeara)
  AAERDATA3[ AAERDATA3$ID %in% l2$ID ,c("DATA24")] <- predict(fit, l2[,c("DATA24","yeara")]) 
}

dataQuality(AAERDATA3$DATA24)
```

The improved data quality for Price at closing is `r round(length(which(AAERDATA3$DATA24!=0))/length(AAERDATA3$DATA24),3)*100`% after missing values were predicted using linear regression.

```{r price2}
#Original data
price_close_mean <- mean(AAERDATA$DATA24)
price_close_count <- length(AAERDATA$DATA24 != 0)
price_close_sum <- sum(AAERDATA$DATA24)
price_close_median <- median(AAERDATA$DATA24)
price_close_sd <- sd(AAERDATA$DATA24)

price_close_mean
price_close_count
price_close_sum
price_close_median
price_close_sd

#Updated data
price_close_mean <- mean(AAERDATA3$DATA24)
price_close_count <- length(AAERDATA3$DATA24)
price_close_sum <- sum(AAERDATA3$DATA24)
price_close_median <- median(AAERDATA3$DATA24)
price_close_sd <- sd(AAERDATA3$DATA24)

price_close_mean
price_close_count
price_close_sum
price_close_median
price_close_sd

plotHistFunc(data.frame(AAERDATA3$DATA24))
```

### Number of Employees

The original data quality for Number of employees is `r round(length(which(AAERDATA3$DATA29!=0))/length(AAERDATA3$DATA29),3)*100` %.

```{r employee1}
dataQuality(AAERDATA3$DATA29)

x <- unique(AAERDATA3[AAERDATA3$DATA29 == 0, c("SMBL")])
AAERDATA4  <- AAERDATA3 %>% select(ID, SMBL, yeara, DATA29) %>% filter(SMBL %in% x)

AAERDATA5 <- AAERDATA4 %>% 
  filter(DATA29 != 0) %>% 
  group_by(SMBL) %>% 
  summarise(n = n()) 

for (i in AAERDATA5[AAERDATA5$n > 2, c("SMBL")])
{
  l1 <- AAERDATA4 %>% filter(SMBL %in% i & DATA29 != 0)
  l2 <- AAERDATA4 %>% filter(SMBL %in% i & DATA29 == 0)
  fit <- lm(data = l1, DATA29~yeara)
  AAERDATA3[ AAERDATA3$ID %in% l2$ID ,c("DATA29")] <- predict(fit, l2[,c("DATA29","yeara")]) 
}

dataQuality(AAERDATA3$DATA2)
```

The improved data quality for Number of employees is `r round(length(which(AAERDATA3$DATA29!=0))/length(AAERDATA3$DATA29),3)*100`% after missing values were predicted using linear regression.

```{r employee2}
#Original data
employees_mean <- mean(AAERDATA$DATA29)
employees_count <- length(AAERDATA$DATA29 != 0)
employees_sum <- sum(AAERDATA$DATA29)
employees_median <- median(AAERDATA$DATA29)
employees_sd <- sd(AAERDATA$DATA29)

employees_mean
employees_count
employees_sum
employees_median
employees_sd

#Updated data
employees_mean <- mean(AAERDATA3$DATA29)
employees_count <- length(AAERDATA3$DATA29)
employees_sum <- sum(AAERDATA3$DATA29)
employees_median <- median(AAERDATA3$DATA29)
employees_sd <- sd(AAERDATA3$DATA29)

employees_mean
employees_count
employees_sum
employees_median
employees_sd

plotHistFunc(data.frame(AAERDATA3$DATA29))
```


```{r}
data_col <- lapply(AAERDATA3, function(x){ length(which(x!=0))/length(x)})
data_col <- data.frame(data_col)
data_col
```

```{r, results="hide"}
replace(AAERDATA3[,  -c(9,17,21,43:45)], TRUE, lapply(AAERDATA3[,  -c(9,17,21,43:45)], NA2mean))
```

```{r}
data_col1 <- lapply(AAERDATA3, function(x){ length(which(x!=0))/length(x)})
data_col1 <- data.frame(data_col1)
data_col1
```

## Identify outliers 

The z-method was selected to identified outlier 

### SALES

```{r}
ggplot(data = AAERDATA, aes(x="Sales", y=DATA12))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Sales") +
  xlab("") +
  ylab("Sales")


AAERDATA_NEW <- data.frame(AAERDATA$DATA12)
colnames(AAERDATA_NEW) <- c("DATA12")

plotHistIQRFunc(AAERDATA_NEW)

lower <- quantile(AAERDATA_NEW$DATA12, .25) - (1.5 * IQR(AAERDATA_NEW$DATA12)) 
upper <- quantile(AAERDATA_NEW$DATA12, .75) + (1.5 * IQR(AAERDATA_NEW$DATA12))
lower
upper


#IQR method
AAERDATA_NEW <- data.frame(AAERDATA_NEW$DATA12[AAERDATA_NEW$DATA12 <= upper])
colnames(AAERDATA_NEW) <- c("DATA12")


ggplot(data = AAERDATA_NEW, aes(x="Sales", y=DATA12))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Sales") +
  xlab("") +
  ylab("Sales")


lower <- quantile(AAERDATA_NEW$DATA12, .25) - 1.5 * IQR(AAERDATA_NEW$DATA12) 
upper <- quantile(AAERDATA_NEW$DATA12, .75) + 1.5 * IQR(AAERDATA_NEW$DATA12)
lower
upper


plotHistIQRFunc(AAERDATA_NEW)
```

```{r sales3}
AAERDATA$DATA12_Z <- scale(AAERDATA$DATA12, center = TRUE, scale = TRUE)

ggplot(data = AAERDATA, aes(x=DATA12_Z))+
  geom_histogram() +
  geom_vline(xintercept = 3, linetype="solid", 
                color = "red", size=1.5) +
  geom_vline(xintercept = -3, linetype="solid", 
                color = "red", size=1.5) 

ggplot(data = AAERDATA, aes(x="Sales", y=DATA12))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Sales") +
  xlab("") +
  ylab("Sales")

AAERDATA_NEW <- data.frame(AAERDATA$DATA12_Z[AAERDATA$DATA12_Z < 3])
colnames(AAERDATA_NEW) <- c("DATA12_Z")
AAERDATA_NEW$DATA12 <- AAERDATA$DATA12[AAERDATA$DATA12_Z < 3]

ggplot(data = AAERDATA_NEW, aes(x=DATA12_Z))+
  geom_histogram() +
  geom_vline(xintercept = 3, linetype="solid", 
                color = "red", size=1.5) +
  geom_vline(xintercept = -3, linetype="solid", 
                color = "red", size=1.5) 

ggplot(data = AAERDATA_NEW, aes(x="Sales", y=DATA12))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Sales") +
  xlab("") +
  ylab("Sales")
```

### Price at closing

```{r price3}
AAERDATA$DATA24_Z <- scale(AAERDATA$DATA24, center = TRUE, scale = TRUE)

ggplot(data = AAERDATA, aes(x=DATA24_Z))+
  geom_histogram() +
  geom_vline(xintercept = 3, linetype="solid", 
                color = "red", size=1.5) +
  geom_vline(xintercept = -3, linetype="solid", 
                color = "red", size=1.5) 

ggplot(data = AAERDATA, aes(x="Price at Closing", y=DATA24))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Price at Closing") +
  xlab("") +
  ylab("Price at Closing")

AAERDATA_NEW <- data.frame(AAERDATA$DATA24_Z[AAERDATA$DATA24_Z < 3])
colnames(AAERDATA_NEW) <- c("DATA24_Z")
AAERDATA_NEW$DATA24 <- AAERDATA$DATA24[AAERDATA$DATA24_Z < 3]

ggplot(data = AAERDATA_NEW, aes(x=DATA24_Z))+
  geom_histogram() +
  geom_vline(xintercept = 3, linetype="solid", 
                color = "red", size=1.5) +
  geom_vline(xintercept = -3, linetype="solid", 
                color = "red", size=1.5)

ggplot(data = AAERDATA_NEW, aes(x="Price at Closing", y=DATA24))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Price at Closing") +
  xlab("") +
  ylab("Price at Closing")
```

### Employees

```{r employee3}
AAERDATA$DATA29_Z <- scale(AAERDATA$DATA29, center = TRUE, scale = TRUE)

ggplot(data = AAERDATA, aes(x=DATA29_Z))+
  geom_histogram() +
  geom_vline(xintercept = 3, linetype="solid", 
                color = "red", size=1.5) +
  geom_vline(xintercept = -3, linetype="solid", 
                color = "red", size=1.5) 

ggplot(data = AAERDATA, aes(x="Employees", y=DATA29))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Employees") +
  xlab("") +
  ylab("Employees")

AAERDATA_NEW <- data.frame(AAERDATA$DATA29_Z[AAERDATA$DATA29_Z < 3])
colnames(AAERDATA_NEW) <- c("DATA29_Z")
AAERDATA_NEW$DATA29 <- AAERDATA$DATA24[AAERDATA$DATA29_Z < 3]


ggplot(data = AAERDATA_NEW, aes(x=DATA29_Z))+
  geom_histogram() +
  geom_vline(xintercept = 3, linetype="solid", 
                color = "red", size=1.5) +
  geom_vline(xintercept = -3, linetype="solid", 
                color = "red", size=1.5)  

ggplot(data = AAERDATA_NEW, aes(x="Employees", y=DATA29))+
  geom_boxplot(outlier.colour="red") +
  ggtitle("Boxplot of Employees") +
  xlab("") +
  ylab("Employees")
```

## Calculate

### Skewness for Sales

```{r skew1}
getskewness(AAERDATA$DATA12)
```

### Skewness for Z-Score standardized SALES

```{r skew2}
getskewness(AAERDATA$DATA12_Z)
```

### Is there evidence of SKEWNESS

There is evidence of skewness for SALES.

## Normal probability plots

### Construct a normal probability plot for Employees

```{r np_employee}
length(which(AAERDATA3$DATA29!=0))/length(AAERDATA3$DATA29)

ggplot(AAERDATA, aes(sample=DATA29)) +
  stat_qq()

qqplot.data(AAERDATA$DATA29)
```

### Derive a new variable – CASH/Total Assets

```{r derive}
length(which(AAERDATA3$DATA1!=0))/length(AAERDATA3$DATA1)
length(which(AAERDATA3$DATA6!=0))/length(AAERDATA3$DATA6)
AAERDATA$cash_asset_ratio <- AAERDATA$DATA1 / AAERDATA$DATA6
```

### Construct a normal probability plot for new Variable.

```{r np_var}
qqplot.data(AAERDATA$cash_asset_ratio)
```

# Part Two

## Read data set

```{r read2}
AAERDATA <- read_excel("~/R/DataMining/Project/AAERDATAHWPart2.xlsx")

#Replace columns names
names(AAERDATA) <- sub("/", "_", names(AAERDATA))
names(AAERDATA) <- sub(" ", "", names(AAERDATA))
names(AAERDATA)[49] <- "LOG_DEFLTA" #69   "LOG(DEFLTA)=LOG(100*(TA_CPI))"
names(AAERDATA)[51] <- "CAL1"       #71   "D60+D181" 
names(AAERDATA)[52] <- "CAL2"       #72   "prc*data25+data6-data60"
names(AAERDATA)[55] <- "Tobin"      #75   "Tobin'sNew" 
names(AAERDATA)[57] <- "MVE_TL"     #77   "MVE_TL"

names(AAERDATA) <- sub("\\(", "_", names(AAERDATA))
names(AAERDATA) <- sub("\\)", "_", names(AAERDATA))
names(AAERDATA) <- sub("=", "", names(AAERDATA))
names(AAERDATA) <- sub("\\+", "_", names(AAERDATA))


AAERDATA1 <- AAERDATA[,c(1:12)]  # categorical

AAERDATA2 <- AAERDATA2A <- AAERDATA[,c(13:73)] # numerical   #13:46, 67:93

AAERDATA2[is.na(AAERDATA2)] <- 0
```

```{r}
summary(AAERDATA1)
AAERDATA1 <- data.frame(lapply(AAERDATA1, function(x) { as.factor(x) }))
```

```{r}
summary(AAERDATA2)
```

## Missing values and erros

1.	Ensure no missing values or errors

This data set had duplicated DATA columns.  Those columns were removed prior to loading to avoid errors during the loading process.  The 

|Row|Original Value|New Value|
|------:|------:|------:|
|1|32416|9/30/1988|
|3|37167|9/30/1988|
|4|37167|10/30/2001|
|5|37167|10/30/2001|

```{r}
AAERDATA2[is.na(AAERDATA2)] <- 0

data_col <- lapply(AAERDATA2, function(x){ length(which(x!=0))/length(x)})
data_col <- data.frame(data_col)

#AAERDATA3 <- AAERDATA2[,data_col[1,] > 0.5]
```

## Standardize

2.	Standardize the data where relevant – ie you don’t standardize GVKEY or SIC codes

```{r}
numeric.index <- unlist(lapply(AAERDATA2, is.numeric)) 

#Normalize
AAERDATA2.Z <- data.frame(lapply(AAERDATA2[,numeric.index], function(x) scale(x, center = FALSE, scale = TRUE)))

colnames(AAERDATA2.Z) <- paste("Z", colnames(AAERDATA2.Z), sep = "_")

plotHistFunc(AAERDATA2.Z)
```

## Variables Types

3.	Which variables are categorical and which are numeric

```{r}
split(names(AAERDATA1),sapply(AAERDATA1, function(x) paste(class(x), collapse=" ")))
```

Continous variables are those with type `numeric`

```{r}
split(names(AAERDATA2),sapply(AAERDATA2, function(x) paste(class(x), collapse=" ")))
```

## Correlation

4.	Examine numeric variables for correlation

```{r}
getCor(AAERDATA2,  val = 0.90)

getCor(AAERDATA2.Z,  val = 0.90)

```

```{r}
# Investigate: PRC (Price – Close) and DATA24 (PriceClose)
# Action: remove DATA24
# Reason: duplicate
AAERDATA2$DATA24 <-  NULL

# Investigate DATA60 (CommonEquityTotal) and DATA216 (StcKhldEqui)
# Common Equity – Total
# This item includes:
#1. Common stock (including effects of common treasury stock)
#2. Capital surplus
#3. Retained earnings
#4. Treasury stock adjustments for both common and nonredeemable preferred stock

# Investigate CAL2 = DATA24 * DATA25 + DATA6 * DATA60 and	MVE_TL = DATA181 + MVEquity

# Investigate CAL1: DATA6 (TotalAssets) + DATA181 (totalLiabi)
# Action: remove DATA6 + DATA181 

AAERDATA2$DATA6 <- NULL

AAERDATA2$DATA181 <- NULL


# Action: remove DATA216
AAERDATA2$DATA216 <-  NULL


# OPINC_TA = DATA178(OperatingIncaftDepr)/DATA6(TotalAssets)	EBIT_TA = EBIt/DATA6(TotalAssets)
# remove DATA178, DATA6 and EBit

AAERDATA2$DATA178 <- NULL

AAERDATA2$EBit <- NULL


# Investigate TotalDebt	= DATA9(LongTermDebtTotal) + DATA34x(DebtinCurrentLiabillities)
# Action: remove DATA9 and DATA34x

AAERDATA2$DATA9 <-  NULL

AAERDATA2$DATA34x <-  NULL
```

## Bar chart (Categorical)
5.	For the categorical variables construct a bar chart

```{r}
plotBarFunc(AAERDATA1[, c(2,6:11)])
```

## Cross tabulation (Categorical)

6.	For each pair of categorical variables construct a cross tabulation

```{r}
nm <- names(AAERDATA1[,c(2,5:7,9:12)])
for (i in nm) {
  print(table(AAERDATA1[,i],AAERDATA1$bktype, dnn=c(i,"bktype")) )
}
```

## Histogram (Numeric)

7.	Construct a histogram of three numeric variables with an overlay of bankruptcy variable (bktype)

```{r}
AAERDATA2$bktype <- AAERDATA1$bktype
nm <- names(AAERDATA2[,1:3])

for (i in seq_along(nm)) {
  p <- ggplot(AAERDATA2,aes_string(x = nm[i], fill = "bktype"),
              aes(y=..density..)) + 
    geom_histogram(alpha = .5) +
    geom_density(alpha=.2, fill="#FF6666")
  print(p)
}
```

## Scatter plot (Numeric)

8.	For three numeric variables construct a scatter plot for each pair.  (Should be three)

```{r}
nm <- c("DATA5","DATA12","DATA25")
for (i in nm) {
  for (j in nm) {
    if (i != j) {
      p <- ggplot(data=AAERDATA2, aes_string(x=i, y=j, col="bktype")) + 
        geom_point() +
        geom_smooth(method = "lm")
       print(p)
    }
  }
}
```

## Correlation analysis

9.	Conduct a correlation analysis of the “Data” variables – exclude the calculated variables

The DATA variables are highly correlated.

```{r}
getCor(AAERDATA2A[,15:33], val = 0.90)
```

## PCA - Numeric data

10.	Generate Principal components based on the numeric data – both DATA and calculated

```{r, warning=FALSE}
AAERDATAS <- data.frame(lapply(AAERDATA2.Z, function(x) scale(x, center = FALSE, scale = TRUE)))
pca <- principal(AAERDATAS,
                  nfactors=ncol(AAERDATAS),
                  rotate="none",
                  scores=TRUE)

pca
```

## PCA components

11.	How many Principal components should you use?

### PCA Results: Eingenvalues and Proportion

From the Eingenvalues and proportion of variance table, 13 components have eingenvalues greater than 1. 

```{r}
pca$values

pca$loadings
```

### PCA Results: Scree plot

Based on scree plot, do not extract more than 13

```{r}
plot(pca$values, type="b",
     main="Scree Plot for AAERDATA Data")
```

### PCA Results: Communality Criterion

Based on communality criterion, extract 13 component if we want to explain at least 85% of the variablity.

```{r}
comm1 <-(loadings(pca)[1,1])^2
comm2 <-comm1+(loadings(pca)[1,2])^2
comm3 <-comm2+(loadings(pca)[1,3])^2
comm4 <-comm3+(loadings(pca)[1,4])^2
comm5 <-comm4+(loadings(pca)[1,5])^2
comm6 <-comm5+(loadings(pca)[1,6])^2
comm7 <-comm6+(loadings(pca)[1,7])^2
comm8 <-comm7+(loadings(pca)[1,8])^2
comm9 <-comm8+(loadings(pca)[1,9])^2
comm10 <-comm9+(loadings(pca)[1,10])^2
comm11 <-comm10+(loadings(pca)[1,11])^2
comm12 <-comm11+(loadings(pca)[1,12])^2
comm13 <-comm12+(loadings(pca)[1,13])^2
comm14 <-comm13+(loadings(pca)[1,14])^2

comm10
comm11
comm12
comm13
comm14
```


## PCA Factor Scores

12.	Plot the factor scores for the number of principal components you identified in #11


## PCA varimax rotation

13.	Apply PCA with varimax rotation for the number of principal components in #10

The number of principal components identified in step #10 is 12.

### Extract 12

```{r, warning=FALSE}
#Varimax rotation
pca.varimax <- principal(AAERDATAS,
                         nfactors=12,
                         rotate="varimax",
                         scores=TRUE)
pca.varimax
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax$scores[,1]
pca_df$PC2 <- pca.varimax$scores[,2]
pca_df$PC3 <- pca.varimax$scores[,3]
pca_df$PC4 <- pca.varimax$scores[,4]
pca_df$PC5 <- pca.varimax$scores[,5]
pca_df$PC6 <- pca.varimax$scores[,6]
pca_df$PC7 <- pca.varimax$scores[,7]
pca_df$PC8 <- pca.varimax$scores[,8]
pca_df$PC9 <- pca.varimax$scores[,9]
pca_df$PC10 <- pca.varimax$scores[,10]
pca_df$PC11 <- pca.varimax$scores[,11]
pca_df$PC12 <- pca.varimax$scores[,12]

pca_glm_12 <- glm(bktype~., data=pca_df)

vif(pca_glm_12)

summary(pca_glm_12)
```

The AIC is `r AIC(pca_glm_12)`.

- PC1 is highly correlated

- PC2 to PC12 are moderately correlated

### Extract 13

```{r, warning=FALSE}
#Varimax rotation
pca.varimax <- principal(AAERDATAS,
                         nfactors=13,
                         rotate="varimax",
                         scores=TRUE)
pca.varimax
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax$scores[,1]
pca_df$PC2 <- pca.varimax$scores[,2]
pca_df$PC3 <- pca.varimax$scores[,3]
pca_df$PC4 <- pca.varimax$scores[,4]
pca_df$PC5 <- pca.varimax$scores[,5]
pca_df$PC6 <- pca.varimax$scores[,6]
pca_df$PC7 <- pca.varimax$scores[,7]
pca_df$PC8 <- pca.varimax$scores[,8]
pca_df$PC9 <- pca.varimax$scores[,9]
pca_df$PC10 <- pca.varimax$scores[,10]
pca_df$PC11 <- pca.varimax$scores[,11]
pca_df$PC12 <- pca.varimax$scores[,12]
pca_df$PC13 <- pca.varimax$scores[,13]

pca_glm_13 <- glm(bktype~., data=pca_df)

vif(pca_glm_13)

summary(pca_glm_13)
```

- PC1 is highly correlated

- PC2 to PC13 are moderately correlated

The AIC is `r AIC(pca_glm_13)`.  The AIC increased.  Extracting 12 components is better since it has the lowerst AIC.

|Component|AIC|Note|
|------:|------:|------:|
|12|`r AIC(pca_glm_12)`|Lowest|
|13|`r AIC(pca_glm_13)`| |

## PCA - Data data only

14.	Examine the difference in PCA results if you use only the Data variables versus using the constructed (calculated variables).

```{r, warning=FALSE}
pca1 <- principal(AAERDATAS[,1:34],
                  nfactors=ncol(AAERDATAS[,1:34]),
                  rotate="none",
                  scores=TRUE)

pca1
```

From the Eingenvalues and proportion of variance table, 8 components have eingenvalues greater than 1. 

```{r}
pca1$values

pca1$loadings
```

Based on scree plot, do not extract more than 8

```{r}
plot(pca1$values, type="b",
     main="Scree Plot for AAERDATA Data - Data Variables")
```

### PCA Results: Communality Criterion

Based on communality criterion, extract 8 component if we want to explain at least 88% of the variablity.

```{r}
comm1 <-(loadings(pca1)[1,1])^2
comm2 <-comm1+(loadings(pca1)[1,2])^2
comm3 <-comm2+(loadings(pca1)[1,3])^2
comm4 <-comm3+(loadings(pca1)[1,4])^2
comm5 <-comm4+(loadings(pca1)[1,5])^2
comm6 <-comm5+(loadings(pca1)[1,6])^2
comm7 <-comm6+(loadings(pca1)[1,7])^2
comm8 <-comm7+(loadings(pca1)[1,8])^2
comm9 <-comm8+(loadings(pca1)[1,9])^2


comm6
comm7
comm8
comm9
```

### Extract 6

```{r, warning=FALSE}
#Varimax rotation
pca.varimax1 <- principal(AAERDATAS[,1:34],
                              nfactors=6,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax1
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax1$scores[,1]
pca_df$PC2 <- pca.varimax1$scores[,2]
pca_df$PC3 <- pca.varimax1$scores[,3]
pca_df$PC4 <- pca.varimax1$scores[,4]
pca_df$PC5 <- pca.varimax1$scores[,5]
pca_df$PC6 <- pca.varimax1$scores[,6]

pca_glm1_6 <- glm(bktype~., data=pca_df)

vif(pca_glm1_6)

summary(pca_glm1_6)
```

The AIC is `r AIC(pca_glm1_6)`.

### Extract 7

```{r, warning=FALSE}
#Varimax rotation
pca.varimax1 <- principal(AAERDATAS[,1:34],
                              nfactors=7,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax1
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax1$scores[,1]
pca_df$PC2 <- pca.varimax1$scores[,2]
pca_df$PC3 <- pca.varimax1$scores[,3]
pca_df$PC4 <- pca.varimax1$scores[,4]
pca_df$PC5 <- pca.varimax1$scores[,5]
pca_df$PC6 <- pca.varimax1$scores[,6]
pca_df$PC7 <- pca.varimax1$scores[,7]


pca_glm1_7 <- glm(bktype~., data=pca_df)

vif(pca_glm1_7)

summary(pca_glm1_7)
```

The AIC is `r AIC(pca_glm1_7)`.

### Extract 8

```{r, warning=FALSE}
#Varimax rotation
pca.varimax1 <- principal(AAERDATAS[,1:34],
                              nfactors=8,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax1
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax1$scores[,1]
pca_df$PC2 <- pca.varimax1$scores[,2]
pca_df$PC3 <- pca.varimax1$scores[,3]
pca_df$PC4 <- pca.varimax1$scores[,4]
pca_df$PC5 <- pca.varimax1$scores[,5]
pca_df$PC6 <- pca.varimax1$scores[,6]
pca_df$PC7 <- pca.varimax1$scores[,7]
pca_df$PC8 <- pca.varimax1$scores[,8]

pca_glm1_8 <- glm(bktype~., data=pca_df)

vif(pca_glm1_8)

summary(pca_glm1_8)
```

The AIC is `r AIC(pca_glm1_8)`.

Extracting 6 components is better since it has the lowerst AIC.

|Component|AIC|Note|
|------:|------:|------:|
|6|`r AIC(pca_glm1_6)`|Lowest|
|7|`r AIC(pca_glm1_7)`||
|8|`r AIC(pca_glm1_8)`||

## PCA - Calculated data only

```{r, warning=FALSE}
pca2 <- principal(AAERDATAS[,35:61],
                  nfactors=ncol(AAERDATAS[,35:61]),
                  rotate="none",
                  scores=TRUE)

pca2
```

From the Eingenvalues and proportion of variance table, 6 components have eingenvalues greater than 1. 

From the proportion of variance table, 7 components have eingenvalues greater than 1. 

```{r}
pca2$values

pca2$loadings
```

Based on scree plot, do not extract more than 7

```{r}
plot(pca2$values, type="b",
     main="Scree Plot for AAERDATA Data - Calculated Variables")
```

### PCA Results: Communality Criterion

Based on communality criterion, extract 7 component if we want to explain at least 87% of the variablity.

```{r}
comm1 <-(loadings(pca1)[1,1])^2
comm2 <-comm1+(loadings(pca1)[1,2])^2
comm3 <-comm2+(loadings(pca1)[1,3])^2
comm4 <-comm3+(loadings(pca1)[1,4])^2
comm5 <-comm4+(loadings(pca1)[1,5])^2
comm6 <-comm5+(loadings(pca1)[1,6])^2
comm7 <-comm6+(loadings(pca1)[1,7])^2
comm8 <-comm7+(loadings(pca1)[1,8])^2

comm5
comm6
comm7
comm8
```

### Extract 5

```{r, warning=FALSE}
#Varimax rotation
pca.varimax2 <- principal(AAERDATAS[,35:61],
                              nfactors=5,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax2
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax2$scores[,1]
pca_df$PC2 <- pca.varimax2$scores[,2]
pca_df$PC3 <- pca.varimax2$scores[,3]
pca_df$PC4 <- pca.varimax2$scores[,4]
pca_df$PC5 <- pca.varimax2$scores[,5]

pca_glm2_5 <- glm(bktype~., data=pca_df)

vif(pca_glm2_5)

summary(pca_glm2_5)
```

The AIC is `r AIC(pca_glm2_5)`.

### Extract 6

```{r, warning=FALSE}
#Varimax rotation
pca.varimax2 <- principal(AAERDATAS[,35:61],
                              nfactors=6,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax2
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax2$scores[,1]
pca_df$PC2 <- pca.varimax2$scores[,2]
pca_df$PC3 <- pca.varimax2$scores[,3]
pca_df$PC4 <- pca.varimax2$scores[,4]
pca_df$PC5 <- pca.varimax2$scores[,5]
pca_df$PC6 <- pca.varimax2$scores[,6]


pca_glm2_6 <- glm(bktype~., data=pca_df)

vif(pca_glm2_6)

summary(pca_glm2_6)
```

The AIC is `r AIC(pca_glm2_6)`.

### Extract 7

```{r, warning=FALSE}
#Varimax rotation
pca.varimax2 <- principal(AAERDATAS[,35:61],
                              nfactors=7,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax2
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax2$scores[,1]
pca_df$PC2 <- pca.varimax2$scores[,2]
pca_df$PC3 <- pca.varimax2$scores[,3]
pca_df$PC4 <- pca.varimax2$scores[,4]
pca_df$PC5 <- pca.varimax2$scores[,5]
pca_df$PC6 <- pca.varimax2$scores[,6]
pca_df$PC7 <- pca.varimax2$scores[,7]


pca_glm2_7 <- glm(bktype~., data=pca_df)

vif(pca_glm2_7)

summary(pca_glm2_7)
```

The AIC is `r AIC(pca_glm2_7)`.

### Extract 8

```{r, warning=FALSE}
#Varimax rotation
pca.varimax2 <- principal(AAERDATAS[,35:61],
                              nfactors=8,
                              rotate="varimax",
                              scores=TRUE)
pca.varimax2
```

```{r}
pca_df <- data.frame(AAERDATA1$bktype)
colnames(pca_df) <- c("bktype")
pca_df$bktype <- as.numeric(pca_df$bktype)
pca_df$PC1 <- pca.varimax2$scores[,1]
pca_df$PC2 <- pca.varimax2$scores[,2]
pca_df$PC3 <- pca.varimax2$scores[,3]
pca_df$PC4 <- pca.varimax2$scores[,4]
pca_df$PC5 <- pca.varimax2$scores[,5]
pca_df$PC6 <- pca.varimax2$scores[,6]
pca_df$PC7 <- pca.varimax2$scores[,7]
pca_df$PC8 <- pca.varimax2$scores[,8]


pca_glm2_8 <- glm(bktype~., data=pca_df)

vif(pca_glm2_8)

summary(pca_glm2_8)
```

The AIC is `r AIC(pca_glm2_8)`.

Extracting 6 components is better since it has the lowerst AIC.

|Component|AIC|Note|
|------:|------:|------:|
|5|`r AIC(pca_glm2_5)`| |
|6|`r AIC(pca_glm2_6)`|Lowest|
|7|`r AIC(pca_glm2_7)`| |
|8|`r AIC(pca_glm2_8)`| |

